Structured Synchronous Reactive Programming for Game Development \\ \Large{Case Study: On Rewriting Pingus from C++ to \CEU}

### 2018-07-12 paper.tex
(>) compile in IEEE
(>) remove two cases, separate listings // now fits in 9 pages

\documentclass[times,twocolumn,final]{elsarticle}

\usepackage{cag}
\usepackage{framed,multirow}

@@

\usepackage[switch,pagewise]{lineno} %Required by command \linenumbers below

\usepackage{xspace}
\usepackage{verbatim}

\usepackage{listings}
\lstdefinelanguage{ceu}{%
  language=C,
  morekeywords={%
    @const, @pure, @safe, CHAN, FOREVER, PAR, PROC, SIGNAL, abort, and,
    await, bool, call, class, data, define, deterministic, do, each, else,
    emit, end, escape, event, every, finalize, hor, implementation, in,
    input, interface, loop, min, native, new, nohold, not, or, output, par,
    pool, pure, return, s, signal, spawn, tag, then, this, traverse, until,
    code, public, private,
    var, watching, when, with},
}
\lstset{
  basicstyle=\small\ttfamily,
  captionpos=b,
  columns=flexible,
  commentstyle=\rmfamily\itshape,
  escapeinside={||},
  frame=tb,
  keepspaces=true,
  keywordstyle=\bfseries,
  language=ceu,
  mathescape=true,
  numbersep=4pt,
  numberstyle=\tiny,
  %upquote=true,
}

@@

\journal{Computers \& Graphics}

@@

\verso{Preprint Submitted for review}

\begin{frontmatter}

@@

\begin{comment}
- fraquezas
    - no quantitative analysis
        - so, only for 2D, non-
- intro a Ceu
    - ssrp and ceu
    - main constructs
    - determinism, termination, cite LCTES, to appear
- acknowledgment
- authors
\end{comment}

@@

%\received{1 February 2017}
\received{\today}
%%%% Do not use the below for submitted manuscripts
%\finalform{28 March 2017}
%\accepted{2 April 2017}
%\availableonline{15 May 2017}
%\communicated{S. Sarkar}

@@

\begin{keyword}
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
%\MSC 41A05\sep 41A10\sep 65D05\sep 65D17
%% Keywords
%\KWD Computers and Graphics\sep Formatting\sep Guidelines
%C++                         \sep
%\CEU                        \sep
Control Flow                \sep
Event-Driven Programming    \sep
Game Logic                  \sep
Synchronous Reactive Programming
\end{keyword}

\end{frontmatter}

\linenumbers

@@

\CEU~\cite{ceu.sensys13,ceu.mod15} is a Esterel-based~\cite{esterel.ieee91}
programming language that originally targets embedded soft real-time systems.
It aims to offer a concurrent, safe, and expressive alternative to C with the
characteristics that follow:
%
\begin{description}
\item [Reactive:] code only executes in reactions to events.
\item [Structured:] programs use structured control mechanisms, such as
    \code{spawn} and \code{await} (to create and suspend activities).
\item [Synchronous:] reactions run atomically and to completion on each line of
    execution, i.e., there's no implicit preemption or real parallelism.
\end{description}
%
Structured reactive programming lets developers write code in direct style,
recovering from the inversion of control imposed by event-driven
execution~\cite{rp.deprecating,rp.rescala,sync_async.cooperative}.
%
%\CEU provides primitives that help describing complex control-flow
%relationships in the game logic more concisely.
%
\CEU supports logical parallelism with a resource-efficient implementation in
terms of memory and CPU usage.
The runtime is single threaded and does not rely on garbage collection for
memory management~\cite{ceu.sensys13}.
%
Existing work in the context of embedded sensor networks evaluates the
expressiveness of \CEU in comparison to event-driven code in C and attests a
reduction in source code size (around 25\%) with a small increase in memory
usage (around 5--10\%)~\cite{ceu.sensys13}.
%
\CEU has also been used in the context of multimedia
systems~\cite{ceumedia.webmedia16} and games~\cite{ceu.mod15}.

@@

\begin{figure*}
%{\footnotesize
\begin{verbatim}
          Path             Ceu     C++   Ceu/C++    Description
          ------------    ----    ----    ----      --------------------------------------
     1    game/           2064    2268    0.91      the main gameplay
     2      ./             710     679    1.05        main functionality
     3      objs/          470     478    0.98        world objects (tiles, traps, etc)
     4      pingu/         884    1111    0.80        pingu behaviors
     5        ./           343     458    0.75          main functionality
     6        actions/     541     653    0.83          pingu actions (bomber, climber, etc)
     7    worldmap/        468     493    0.95      campaign worldmap
     8    screens/        1109    1328    0.84      menus and screens
     9        option/      347     357    0.97        option menu
    10        others/      762     971    0.78        other menus and screens
    11    misc/             56      46    1.22      miscellaneous functionality
                          ----    ----    ----
                          3697    4135    0.89
\end{verbatim}
%}
\caption{The Pingus codebase directory tree.
\label{tab.tree}
}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}[h]{0.50\linewidth}
\begin{lstlisting}[numbers=right]
ArmageddonButton::ArmageddonButton(<...>):
    RectComponent(<...>),
    pressed(false); // button is not initially pressed
    press_time(0);  // how long since 1st click?
    <...>
{
    <...>
}

@@

void ArmageddonButton::on_click (<...>) {
    if (pressed) {
      send_armageddon_event();
    } else {
      pressed = true;
    }
}
\end{lstlisting}
\centering\small{\ax Implementation in C++}
\end{minipage}%
%
\begin{minipage}[h]{0.50\linewidth}
%\begin{lstlisting}
\begin{lstlisting}[xleftmargin=2em]
do
    var RectComponent but = <...>;
    <...>
    loop do
      await but.on_click;
      watching 1s do
        await but.on_click;
        break;
      end
    end
    <...>
    emit game.armageddon;
end

.
\end{lstlisting}
\centering\small{\bx Implementation in \CEU}
\end{minipage}%
%\rule{8.4cm}{0.37pt}
\caption{ Detecting double-clicks in the \emph{Armageddon button}.
\label{lst.armageddon}
}
\end{figure*}

@@

\subsubsection{Case Study: The \emph{Bomber Action} Animation Sequence}
\label{sec.pats.fsms.2}

@@

\begin{figure*}
\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[numbers=right]
Bomber::Bomber (Pingu* p) :
    <...>
    spr(<...>),         // bomber sprite
    sound_ok(false),    // tracks state 2
    particle_ok(false), // tracks state 3
    colmap_ok(false),   // tracks state 3
    gfx_ok(false)       // tracks state 4
{
    <...>
    // 1. plays a "Oh no!" sound.
    play_sound("ohno");
}

@@

void Bomber::draw (SceneContext& gc) {
    // 3. particles, terrain, explosion sprite
    // 4. tick: hides the explosion sprite
    if (spr.frame()==13 && !gfx_ok) {
      gfx_ok = true;
      gc.color().draw(explo_surf, <...>);
    }
    gc.color().draw(spr, pingu->get_pos());
}
\end{lstlisting}
\centering\small{\ax Implementation in C++}
\end{minipage}
%
\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[xleftmargin=2em]
code/await Bomber (void) -> ActionName
do
    <...>
    spawn Mover(); // movement in background
    var Sprite spr = spawn Sprite(<...>);
                 // frame animation in background
    watching spr do
      // 1. plays a "Oh no!" sound.
      {play_sound("ohno")};

@@

    // 5. kills the pingu
    escape DEAD;
end

.
\end{lstlisting}
\centering\small{\bx Implementation in \CEU}
\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ The \emph{Bomber action} sequence.
\label{lst.bomber}
}
\end{figure*}

@@

The equivalent code in \CEU for the \emph{Bomber action} in
Figure~\ref{lst.bomber}.b does not rely on state variables and reflects the
sequential state machine implicitly, using \code{await} statements to separate
the effects in direct style.
%
The \code{Bomber} is a \code{code/await} abstraction of \CEU, which is similar
to a coroutine or fiber~\cite{sync_async.cooperative}: a subroutine that
retains runtime state, such as local variables and the program counter, across
reactions to events (i.e., across \code{await} statements).
The pingu movement and sprite animation are isolated in two other
\code{code/await} abstractions and execute in separate through the \code{spawn}
primitive (ln 4--5).
%In \CEU, if multiple abstractions react to the same event, the scheduler
%employs lexical order to preserve determinism, i.e., the \code{spawn} that
%appears first in the source code reacts first~\cite{ceu.mod15}.
The event \code{game.update} (ln 12,16,24) is analogous to the \code{update}
callback of C++ and occurs on every game frame.
%
The code tracks the animation aliveness (ln 7--27) and, on termination,
performs the last bomber effect, killing the pingu (ln 30).
As soon as the animation starts, the code performs the first effect (ln 9).
The intermediate effects are performed when the corresponding conditions occur
(ln 12,16,24).
The \code{do-end} block (ln 19--25), restricts the lifespan of the
single-frame explosion sprite (ln 21): after the next game tick (ln 24), the
block terminates and automatically destroys the spawned abstraction (removing
it from the screen).
%
In contrast with the implementation in C++, all effects occur in a contiguous
chunk of code (ln 7--30), which handles no extra functionality.

@@

\begin{figure*}
\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[numbers=right]
StoryDot::StoryDot(FileReader& reader) :
    show_credits(false), // do not show by default
{
    <...>
    reader.read("credits", show_credits);
}             // from file

@@

///

@@

void StoryScreenComp::next_text() {
    if (!displayed) {
      <...>
    } else {
      <...>
      if (!pages.empty()) {
        <...>
      } else {
        if (show_credits) {
          replace(<Credits>(<...>));
        } else {
          pop();
        }
      }
    }
}
\end{lstlisting}
\centering\small{\ax Implementation in C++}
\end{minipage}
%
\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[xleftmargin=2em]
loop do
    var int ret = await Worldmap();
    if ret==CREDITS or ret==BACK then
      <...>
      var bool is_click = await Story();
      if is_click and ret==CREDITS then
        <...>
        await Credits();
      end
    else
      <...>
    end
end

.
\end{lstlisting}
\centering\small{\bx Implementation in \CEU}
\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ Transition from \emph{Story} to \emph{Credits screen}.
\label{lst.story}
}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[numbers=right]
class Bomber : public Action {
    <...>
    Sprite sprite;
}

@@

void Bomber::draw () {
    <...>
    sprite.draw();
}
\end{lstlisting}
\centering\small{\ax Implementation in C++}
\end{minipage}
%
\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[xleftmargin=2em]
code/await Bomber (void) -> ActionName do
    <...>
    var Sprite sprite = spawn Sprite(<...>);
    <...>
end

.
\end{lstlisting}
\centering\small{\bx Implementation in \CEU}
\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ \emph{Bomber action} \code{draw} and \code{update} dispatching.
\label{lst.hier}
}
\end{figure*}

@@

\subsubsection{Case Study: Static Game UI Widgets}

@@

In \CEU, the UI entities that coexist are simply created in the same lexical
block of the \code{Game} abstraction in Figure~\ref{lst.ui}.b (ln 3--5).
%
Since abstractions can react independently, they do not require a dispatching
container.
%
Lexical lifespan never requires containers, allocation and deallocation, or
explicit references.
In addition, all required memory is known at compile time, similarly to
stack-allocated local variables.
%
The \emph{Bomber action} of Section~\ref{sec.pats.fsms.2} also relies on
lexical scope to delimit the lifespan of the explosion sprite into a single
frame (Figure~\ref{lst.bomber}, ln 19--25).

@@

\newpage
\bibliographystyle{cag-num-names}
\bibliography{my,other}

>>>>

\documentclass[10pt, conference, compsocconf]{IEEEtran}
%\documentclass[times,twocolumn,final]{elsarticle}

%\usepackage{cag}
%\usepackage{framed,multirow}

\usepackage{cite}

@@

\usepackage{xspace}
\usepackage{verbatim}
\usepackage[pdftex]{graphicx}

\usepackage{listings}
\lstdefinelanguage{ceu}{%
  language=C,
  morekeywords={%
    @const, @pure, @safe, CHAN, FOREVER, PAR, PROC, SIGNAL, abort, and,
    await, bool, call, class, data, define, deterministic, do, each, else,
    emit, end, escape, event, every, finalize, hor, implementation, in,
    input, interface, loop, min, native, new, nohold, not, or, output, par,
    pool, pure, return, s, signal, spawn, tag, then, this, traverse, until,
    code, public, private,
    var, watching, when, with},
}
\lstset{
  basicstyle=\footnotesize\ttfamily,
  captionpos=b,
  columns=flexible,
  commentstyle=\rmfamily\itshape,
  escapeinside={||},
  frame=tb,
  keepspaces=true,
  keywordstyle=\bfseries,
  language=ceu,
  mathescape=true,
  numbersep=4pt,
  numberstyle=\tiny,
  %upquote=true,
}

@@

%\journal{Computers \& Graphics}

@@

%\verso{Preprint Submitted for review}

%\begin{frontmatter}

@@

\begin{comment}
- fraquezas
    - no quantitative analysis
        - so, only for 2D, non-
- intro a Ceu
    - ssrp and ceu
    - main constructs
    - determinism, termination, cite LCTES, to appear
- acknowledgment
- authors

- garantias
- na intro, falar pq apply to most games, quais são e como e óbivo que se aplicam

- similar metrics in previous work
    - state machines

- inverter related work?

- qual o custo de CPU p/ o pingus?
    - falar do sensys
    - comentar sobre AAA, talvez para partes do jogo?

- brief intro about SSRP, Ceu

- drawbacks de SSRP?

- on_click / updated
    - different methods

Ex.: Fig.3.b): how is but.on_click updated? How does the "emit game.armagedon" arrives to pingus?
Using events? (please see below).
Fig.7.b): how is variable game.update modified?

Communication in Ceu requires more detail. Pg. 10 states that events are broadcasted. Broadcast is the core
of the observer pattern and it should be simple to achieve the same effect in OOP. Nevertheless, broadcast communication
can be ineffective and make applications very difficult to understand/debug. The limitations of broadcast should
be discussed in detail. If events are broadcasted, are filters supported?

\end{comment}

@@

%\received{1 February 2017}
%\received{\today}
%%%% Do not use the below for submitted manuscripts
%\finalform{28 March 2017}
%\accepted{2 April 2017}
%\availableonline{15 May 2017}
%\communicated{S. Sarkar}

\maketitle

@@

\begin{IEEEkeywords}
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
%\MSC 41A05\sep 41A10\sep 65D05\sep 65D17
%% Keywords
%\KWD Computers and Graphics\sep Formatting\sep Guidelines
%C++                         \sep
%\CEU                        \sep
Control Flow                     ;
Event-Driven Programming         ;
Game Logic                       ;
Synchronous Reactive Programming ;
\end{IEEEkeywords}

%\end{frontmatter}

%\linenumbers

@@

\CEU~\cite{ceu.sensys13,ceu.mod15} is a Esterel-based~\cite{esterel.ieee91}
programming language that originally targets embedded soft real-time systems.
It aims to offer a concurrent, safe, and expressive alternative to C with the
characteristics that follow:
%
\begin{itemize}
\item \textbf{Reactive:} code only executes in reactions to events.
\item \textbf{Structured:} programs use structured control mechanisms, such as
    \code{spawn} and \code{await} (to create and suspend activities).
\item \textbf{Synchronous:} reactions run atomically and to completion on each line of
    execution, i.e., there's no implicit preemption or real parallelism.
\end{itemize}
%
Structured reactive programming lets developers write code in direct style,
recovering from the inversion of control imposed by event-driven
execution~\cite{rp.deprecating,rp.rescala,sync_async.cooperative}.
%
%\CEU provides primitives that help describing complex control-flow
%relationships in the game logic more concisely.
%
\CEU supports logical parallelism with a resource-efficient implementation in
terms of memory and CPU usage.
The runtime is single threaded and does not rely on garbage collection for
memory management~\cite{ceu.sensys13}.
%
Existing work in the context of embedded sensor networks evaluates the
expressiveness of \CEU in comparison to event-driven code in C and attests a
reduction in source code size (around 25\%) with a small increase in memory
usage (around 5--10\%)~\cite{ceu.sensys13}.
%
\CEU has also been used in the context of multimedia
systems~\cite{ceumedia.webmedia16} and games~\cite{ceu.mod15}.

@@

\begin{figure*}
{\footnotesize
\begin{verbatim}
          Path             Ceu     C++   Ceu/C++    Description
          ------------    ----    ----    ----      --------------------------------------
     1    game/           2064    2268    0.91      the main gameplay
     2      ./             710     679    1.05        main functionality
     3      objs/          470     478    0.98        world objects (tiles, traps, etc)
     4      pingu/         884    1111    0.80        pingu behaviors
     5        ./           343     458    0.75          main functionality
     6        actions/     541     653    0.83          pingu actions (bomber, climber, etc)
     7    worldmap/        468     493    0.95      campaign worldmap
     8    screens/        1109    1328    0.84      menus and screens
     9        option/      347     357    0.97        option menu
    10        others/      762     971    0.78        other menus and screens
    11    misc/             56      46    1.22      miscellaneous functionality
                          ----    ----    ----
                          3697    4135    0.89
\end{verbatim}
}
\caption{The Pingus codebase directory tree.
\label{tab.tree}
}
\end{figure*}

@@

\begin{figure}[t]
%\begin{minipage}[h]{0.50\linewidth}
\begin{lstlisting}[numbers=right]
ArmageddonButton::ArmageddonButton(<...>):
    RectComponent(<...>),
    pressed(false); // button is not initially pressed
    press_time(0);  // how long since 1st click?
    <...>
{
    <...>
}

@@

void ArmageddonButton::on_click (<...>) {
    if (pressed) {
      send_armageddon_event();
    } else {
      pressed = true;
    }
}
\end{lstlisting}
%\centering\small{\ax Implementation in C++}
%\end{minipage}%
%
%\begin{minipage}[h]{0.50\linewidth}
%\begin{lstlisting}
\end{figure}
\begin{figure}[t]
\begin{lstlisting}[xleftmargin=2em]
do
    var RectComponent but = <...>;
    <...>
    loop do
      await but.on_click;
      watching 1s do
        await but.on_click;
        break;
      end
    end
    <...>
    emit game.armageddon;
end
\end{lstlisting}
%\centering\small{\bx Implementation in \CEU}
%\end{minipage}%
%\rule{8.4cm}{0.37pt}
\caption{ Detecting double-clicks in the \emph{Armageddon button}.
\label{lst.armageddon}
}
\end{figure}

@@

\begin{comment}
\subsubsection{Case Study: The \emph{Bomber Action} Animation Sequence}
\label{sec.pats.fsms.2}

@@

\begin{figure}[t]
%\begin{minipage}[t]{0.50\linewidth}
%\begin{lstlisting}[numbers=right]
Bomber::Bomber (Pingu* p) :
    <...>
    spr(<...>),         // bomber sprite
    sound_ok(false),    // tracks state 2
    particle_ok(false), // tracks state 3
    colmap_ok(false),   // tracks state 3
    gfx_ok(false)       // tracks state 4
{
    <...>
    // 1. plays a "Oh no!" sound.
    play_sound("ohno");
}

@@

void Bomber::draw (SceneContext& gc) {
    // 3. particles, terrain, explosion sprite
    // 4. tick: hides the explosion sprite
    if (spr.frame()==13 && !gfx_ok) {
      gfx_ok = true;
      gc.color().draw(explo_surf, <...>);
    }
    gc.color().draw(spr, pingu->get_pos());
}
%\end{lstlisting}
%\centering\small{\ax Implementation in C++}
%\end{minipage}
\end{figure}
%
\begin{figure}[t]
%\begin{minipage}[t]{0.50\linewidth}
%\begin{lstlisting}[xleftmargin=2em]
code/await Bomber (void) -> ActionName
do
    <...>
    spawn Mover(); // movement in background
    var Sprite spr = spawn Sprite(<...>);
                 // frame animation in background
    watching spr do
      // 1. plays a "Oh no!" sound.
      {play_sound("ohno")};

@@

    // 5. kills the pingu
    escape DEAD;
end
\end{lstlisting}
%\centering\small{\bx Implementation in \CEU}
%\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ The \emph{Bomber action} sequence.
\label{lst.bomber}
}
\end{figure}

@@

The equivalent code in \CEU for the \emph{Bomber action} in
Figure~\ref{lst.bomber}.b does not rely on state variables and reflects the
sequential state machine implicitly, using \code{await} statements to separate
the effects in direct style.
%
The \code{Bomber} is a \code{code/await} abstraction of \CEU, which is similar
to a coroutine or fiber~\cite{sync_async.cooperative}: a subroutine that
retains runtime state, such as local variables and the program counter, across
reactions to events (i.e., across \code{await} statements).
The pingu movement and sprite animation are isolated in two other
\code{code/await} abstractions and execute in separate through the \code{spawn}
primitive (ln 4--5).
%In \CEU, if multiple abstractions react to the same event, the scheduler
%employs lexical order to preserve determinism, i.e., the \code{spawn} that
%appears first in the source code reacts first~\cite{ceu.mod15}.
The event \code{game.update} (ln 12,16,24) is analogous to the \code{update}
callback of C++ and occurs on every game frame.
%
The code tracks the animation aliveness (ln 7--27) and, on termination,
performs the last bomber effect, killing the pingu (ln 30).
As soon as the animation starts, the code performs the first effect (ln 9).
The intermediate effects are performed when the corresponding conditions occur
(ln 12,16,24).
The \code{do-end} block (ln 19--25), restricts the lifespan of the
single-frame explosion sprite (ln 21): after the next game tick (ln 24), the
block terminates and automatically destroys the spawned abstraction (removing
it from the screen).
%
In contrast with the implementation in C++, all effects occur in a contiguous
chunk of code (ln 7--30), which handles no extra functionality.
\end{comment}

@@

\begin{figure}[t]
%\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[numbers=right]
StoryDot::StoryDot(FileReader& reader) :
    show_credits(false), // do not show by default
{
    <...>
    reader.read("credits", show_credits);
}             // from file

@@

/// class separator ///

@@

void StoryScreenComp::next_text() {
    if (!displayed) {
      <...>
    } else {
      <...>
      if (!pages.empty()) {
        <...>
      } else {
        if (show_credits) {
          replace(<Credits>(<...>));
        } else {
          pop();
        }
      }
    }
}
%\end{lstlisting}
%\centering\small{\ax Implementation in C++}
%\end{minipage}
\end{figure}
%
\begin{figure}[t]
%\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[xleftmargin=2em]
loop do
    var int ret = await Worldmap();
    if ret==CREDITS or ret==BACK then
      <...>
      var bool is_click = await Story();
      if is_click and ret==CREDITS then
        <...>
        await Credits();
      end
    else
      <...>
    end
end
\end{lstlisting}
%\centering\small{\bx Implementation in \CEU}
%\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ Transition from \emph{Story} to \emph{Credits screen}.
\label{lst.story}
}
\end{figure}

@@

\begin{figure}[t]
%\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[numbers=right]
class Bomber : public Action {
    <...>
    Sprite sprite;
}

@@

void Bomber::draw () {
    <...>
    sprite.draw();
}
\end{lstlisting}
%\centering\small{\ax Implementation in C++}
%\end{minipage}
\end{figure}
%
\begin{figure}[t]
%\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[xleftmargin=2em]
code/await Bomber (void) -> ActionName do
    <...>
    var Sprite sprite = spawn Sprite(<...>);
    <...>
end
\end{lstlisting}
%\centering\small{\bx Implementation in \CEU}
%\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ \emph{Bomber action} \code{draw} and \code{update} dispatching.
\label{lst.hier}
}
\end{figure}

@@

\begin{comment}
\subsubsection{Case Study: Static Game UI Widgets}

@@

In \CEU, the UI entities that coexist are simply created in the same lexical
block of the \code{Game} abstraction in Figure~\ref{lst.ui}.b (ln 3--5).
%
Since abstractions can react independently, they do not require a dispatching
container.
%
Lexical lifespan never requires containers, allocation and deallocation, or
explicit references.
In addition, all required memory is known at compile time, similarly to
stack-allocated local variables.
%
The \emph{Bomber action} of Section~\ref{sec.pats.fsms.2} also relies on
lexical scope to delimit the lifespan of the explosion sprite into a single
frame (Figure~\ref{lst.bomber}, ln 19--25).
\end{comment}

@@

%\newpage
\bibliographystyle{abbrv}
\bibliography{my,other}
